---
title: "Simulation: parameter recovery as a function of available data"
subtitle: "Idiographic Memory Modelling in ACT-R using Alternating Maximum Likelihood Estimation"
author: "Thomas Wilschut"
date: "Last updated: `r Sys.Date()`"
output:
  github_document:
    toc: yes
---

This markdown file is can be used to reproduce the results reported in the paper: 
Idiographic Memory Modelling in ACT-R using Alternating Maximum Likelihood Estimation

This script is used for the recovery analysis, reported in the section:
Simulation: Recovering parameters from synthetic data.

# Setup

```{r setup, message=FALSE, warning=FALSE}
library(data.table)
library(here)
library(ggplot2)
library(furrr)
library(patchwork)
library(ggsci)
library(Rcpp)
library(GGally)
library(tidyverse)

# Parallel backend
plan(multisession, workers = availableCores() - 1)

# Source simulation + fitting functions (which also compile actr_ll.cpp)
source(here::here("R", "sim-mle.R"))

# Plot aesthetics
col_blue <- "#0571b0"
col_red  <- "#ca0020"
plottheme <- theme_bw(base_size = 12) 
```



# Simulate a larger dataset for recovery

For the recovery analysis we generate a single large dataset (50 repetitions per fact) once, and then subset it down to smaller numbers of repetitions and facts in the grid below. Generating the maximum amount of data up front and subsetting it keeps the underlying learners and facts identical across all data-quantity conditions, so differences in recovery reflect the amount of data rather than a different random draw.

```{r}
sim <- simulate_fact_learner_data(
  n_learners    = 50,
  n_facts       = 15,
  n_repetitions = 50,   # generate plenty; we'll subset later
  seed          = 999
)

sim_test <- sim[is_study == FALSE]
```

# Run recovery across grid

We vary the amount of available data along two axes: the number of repetitions per fact (holding facts fixed at 15) and the number of facts per learner (holding repetitions fixed at 10). For each grid point we refit all 50 learners and record the recovered parameters. As with the greedy step, this is expensive, so the chunk is `eval = FALSE` and the results are cached to `data/processed/fits/`; the next chunk reloads them.

```{r, eval = F}
rep_levels  <- c(1, 2, 4, 6, 8, 10, 15, 20, 30)
fact_levels <- c(1, 2, 4, 6, 8, 10, 15, 20, 30)

fits_vary_reps <- 
  map_dfr(rep_levels, function(nr) {
    cat(sprintf("\n=== Fitting n_reps = %d ===\n", nr))
    res <- run_recovery_grid(sim, n_reps = nr, n_facts = 15)
    res$n_repetitions <- nr
    res$n_facts_grid  <- 15
    res
  })

fits_vary_facts <- 
  map_dfr(fact_levels, function(nf) {
    cat(sprintf("\n=== Fitting n_facts = %d ===\n", nf))
    res <- run_recovery_grid(sim, n_reps = 10, n_facts = nf)
    res$n_repetitions <- 10
    res$n_facts_grid  <- nf
    res
  })

saveRDS(fits_vary_reps,  here::here("data", "processed", "fits", "results_recovery_vary_reps.rds"))
saveRDS(fits_vary_facts, here::here("data", "processed", "fits", "results_recovery_vary_facts.rds"))
```

# Load results (if saved earlier)

```{r}
fits_vary_reps  <- readRDS(here::here("data", "processed", "fits", "results_recovery_vary_reps.rds"))
fits_vary_facts <- readRDS(here::here("data", "processed", "fits", "results_recovery_vary_facts.rds"))

names(fits_vary_reps)
```

# Merge true parameters

To assess recovery we need the ground-truth parameter values alongside the recovered estimates. Here we extract the true person-level parameters (one row per learner) and join them onto the recovered fits from both grids.

```{r}
true_params <- sim_test[, .(
  true_alpha = true_alpha[1],
  true_tau   = true_tau[1],
  true_s     = true_s[1],
  true_F     = true_lf[1],
  true_ter   = true_ter[1]
), by = user_id]

recovery_reps  <- merge(fits_vary_reps,  true_params, by = "user_id")
recovery_facts <- merge(fits_vary_facts, true_params, by = "user_id")
```

# Recovery (Figure 5)

This builds the combined recovery figure (Figure 5 in the paper). Panel A shows true-vs-recovered scatter plots at the reference data quantity (10 repetitions, 15 facts); panels B–E show how recovery (correlation) and mean absolute error change as a function of repetitions per fact and number of facts, with the reference condition highlighted; panel F shows the pairwise correlations between the recovered estimates, which reveal the trade-offs between parameters with partly overlapping behavioural effects. The helper functions `make_cor_summary`/`make_error_summary` and `plot_cor`/`plot_error` are defined here to avoid repeating the same summarising and plotting code for each panel.

```{r}
# ── Shared aesthetics (consistent with other plots) ───────────────────────────
param_levels <- c(
  "Speed of Forgetting", "Retrieval threshold", "Activation noise",
  "Latency factor", "Non-retrieval time"
)

param_colors    <- setNames(pal_aaas()(5), param_levels)
param_shapes    <- setNames(c(16, 17, 15, 18, 8),                              param_levels)
param_linetypes <- setNames(c("solid", "dashed", "dotdash", "longdash", "twodash"), param_levels)

# ── Panel A: Scatter plots ────────────────────────────────────────────────────
subset_10 <- recovery_reps[n_repetitions == 10]

plot_df <- subset_10 %>%
  pivot_longer(cols = c(alpha, tau, s, lf, ter),
               names_to = "param", values_to = "estimate") %>%
  mutate(
    true_value = case_when(
      param == "alpha" ~ true_alpha,
      param == "tau"   ~ true_tau,
      param == "s"     ~ true_s,
      param == "lf"    ~ true_F,
      param == "ter"   ~ true_ter
    ),
    param_label = recode(param,
      "alpha" = "Speed of Forgetting",
      "tau"   = "Retrieval threshold",
      "s"     = "Activation noise",
      "lf"    = "Latency factor",
      "ter"   = "Non-retrieval time"
    ),
    param_label = factor(param_label, levels = param_levels)
  )

learner_corrs <- plot_df %>%
  group_by(param_label) %>%
  summarise(r = cor(estimate, true_value, use = "complete.obs"), .groups = "drop") %>%
  mutate(r_lab = paste0("r = ", format(round(r, 2), nsmall = 2)))

p_recovery <- ggplot(plot_df, aes(x = true_value, y = estimate, color = param_label)) +
  geom_point(alpha = 0.85, size = 2) +
  geom_smooth(method = "lm", se = TRUE, colour = "black") +
  geom_text(data = learner_corrs,
            aes(x = -Inf, y = Inf, label = r_lab),
            inherit.aes = FALSE, hjust = -0.1, vjust = 2, size = 4) +
  facet_wrap(~ param_label, scales = "free", ncol = 5) +
  coord_cartesian(clip = "off") +
  scale_color_manual(values = param_colors) +
  plottheme +
  theme(
    legend.position  = "none",
    strip.background = element_rect(fill = "grey85", colour = "grey60"),
    strip.text       = element_text(size = 9)
  ) +
  labs(x = "Original value", y = "Recovered value")

# ── Helper functions ──────────────────────────────────────────────────────────
make_cor_summary <- function(df, group_var) {
  df %>%
    group_by(!!sym(group_var)) %>%
    summarise(
      cor_alpha = cor(true_alpha, alpha, use = "complete.obs"),
      cor_tau   = cor(true_tau,   tau,   use = "complete.obs"),
      cor_s     = cor(true_s,     s,     use = "complete.obs"),
      cor_F     = cor(true_F,     lf,    use = "complete.obs"),
      cor_ter   = cor(true_ter,   ter,   use = "complete.obs"),
      .groups = "drop"
    ) %>%
    pivot_longer(starts_with("cor_"), names_to = "parameter", values_to = "correlation") %>%
    mutate(
      parameter   = sub("cor_", "", parameter),
      param_label = recode(parameter,
        "alpha" = "Speed of Forgetting",
        "tau"   = "Retrieval threshold",
        "s"     = "Activation noise",
        "F"     = "Latency factor",
        "ter"   = "Non-retrieval time"
      ),
      param_label = factor(param_label, levels = param_levels)
    )
}

make_error_summary <- function(df, group_var) {
  df %>%
    mutate(
      alpha_error = abs(true_alpha - alpha),
      tau_error   = abs(true_tau   - tau),
      s_error     = abs(true_s     - s),
      F_error     = abs(true_F     - lf),
      ter_error   = abs(true_ter   - ter)
    ) %>%
    pivot_longer(ends_with("_error"), names_to = "parameter", values_to = "abs_error") %>%
    group_by(parameter, !!sym(group_var)) %>%
    summarise(mean_error = mean(abs_error, na.rm = TRUE),
              se_error   = sd(abs_error,   na.rm = TRUE) / sqrt(n()),
              .groups = "drop") %>%
    mutate(
      param_label = recode(parameter,
        "alpha_error" = "Speed of Forgetting",
        "tau_error"   = "Retrieval threshold",
        "s_error"     = "Activation noise",
        "F_error"     = "Latency factor",
        "ter_error"   = "Non-retrieval time"
      ),
      param_label = factor(param_label, levels = param_levels)
    )
}

# ── Shared scale layers ───────────────────────────────────────────────────────
scales_shared <- list(
  scale_color_manual(values = param_colors),
  scale_linetype_manual(values = param_linetypes),
  scale_shape_manual(values = param_shapes),
  guides(colour = guide_legend(override.aes = list(
    linetype = param_linetypes, shape = param_shapes)),
    linetype = "none", shape = "none")
)

plot_cor <- function(df, x_var, x_lab, highlight_x = NULL) {
  p <- ggplot(df, aes(x = factor(!!sym(x_var)), y = correlation,
                      colour = param_label, linetype = param_label,
                      shape = param_label, group = param_label))
  if (!is.null(highlight_x)) {
    p <- p + annotate("rect",
      xmin = which(levels(factor(df[[x_var]])) == as.character(highlight_x)) - 0.4,
      xmax = which(levels(factor(df[[x_var]])) == as.character(highlight_x)) + 0.4,
      ymin = -Inf, ymax = Inf, fill = "grey80", alpha = 0.4)
  }
  p + geom_line(linewidth = 1) + geom_point(size = 2) +
    scale_y_continuous(limits = c(0, 1)) +
    scales_shared +
    plottheme +
    theme(legend.position = "none") +
    labs(x = x_lab, y = "Recovery (correlation r)")
}

plot_error <- function(df, x_var, x_lab, highlight_x = NULL) {
  p <- ggplot(df, aes(x = factor(!!sym(x_var)), y = mean_error,
                      colour = param_label, linetype = param_label,
                      shape = param_label, group = param_label))
  if (!is.null(highlight_x)) {
    p <- p + annotate("rect",
      xmin = which(levels(factor(df[[x_var]])) == as.character(highlight_x)) - 0.4,
      xmax = which(levels(factor(df[[x_var]])) == as.character(highlight_x)) + 0.4,
      ymin = -Inf, ymax = Inf, fill = "grey80", alpha = 0.4)
  }
  p + geom_line(linewidth = 1) + geom_point(size = 2.5) +
    geom_errorbar(aes(ymin = mean_error - se_error, ymax = mean_error + se_error),
                  width = 0.3, alpha = 0.6) +
    scales_shared +
    plottheme +
    theme(legend.position = "none") +
    labs(x = x_lab, y = "Mean absolute error")
}

# ── Panels B–E ────────────────────────────────────────────────────────────────
cor_reps   <- make_cor_summary(recovery_reps, "n_repetitions")
error_reps <- make_error_summary(recovery_reps, "n_repetitions")
cor_facts  <- make_cor_summary(recovery_facts, "n_facts_grid")
error_facts <- make_error_summary(recovery_facts, "n_facts_grid")

p_corr_reps  <- plot_cor(cor_reps  %>% filter(n_repetitions < 20), "n_repetitions", "Repetitions per fact (15 facts)", highlight_x = 10)
p_error_reps <- plot_error(error_reps %>% filter(n_repetitions < 20), "n_repetitions", "Repetitions per fact (15 facts)", highlight_x = 10)
p_corr_facts  <- plot_cor(cor_facts,  "n_facts_grid", "N facts (10 repetitions)", highlight_x = 15)
p_error_facts <- plot_error(error_facts, "n_facts_grid", "N facts (10 repetitions)", highlight_x = 15)

# ── Panel F: Pairwise correlations ────────────────────────────────────────────
pairs_df <- subset_10[, .(alpha, tau, s, lf, ter)]


pairs_plot <- ggpairs(
  as.data.frame(pairs_df),
  upper = list(continuous = function(data, mapping, ...) {
    ggally_cor(data, mapping, ...)
  }),
  lower = list(continuous = function(data, mapping, ...) {
    ggplot(data, mapping) +
      geom_point(alpha = 0.3) +
      geom_smooth(method = "lm", se = TRUE, color = "black")
  }),
  columnLabels = c("Speed of\nForgetting", "Retrieval\nthreshold",
                 "Activation\nnoise", "Latency\nfactor", "Non-retrieval\ntime")
) +
  plottheme +
  theme(strip.background = element_rect(fill = "grey85", colour = "grey60"),
        strip.text       = element_text(size = 8),
        axis.text.x      = element_text(angle = 90, hjust = 1))


# ── Combined figure ───────────────────────────────────────────────────────────
pairs_grob <- GGally::ggmatrix_gtable(pairs_plot)

combined <-
  (p_recovery) /
  ((p_corr_reps + p_error_reps) / (p_corr_facts + p_error_facts) |
     wrap_elements(pairs_grob)) +
  plot_layout(heights = c(0.8, 2), widths = c(1.8, 1)) +
  plot_annotation(tag_levels = "A") &
  theme(plot.tag = element_text(face = "bold", size = 14))

combined
# Paper Figure 5: recovery of participant-level parameters as a function of available data
ggsave(here::here("output", "recovery_combined.png"), combined, width = 11, height = 9)
```


# Fact-level Δα recovery (Figure 6)

Besides the person-level parameters, the model estimates a Speed of Forgetting offset (Δα) for each individual fact. Here we assess how well those fact-level offsets are recovered (Figure 6 in the paper): panel A is a true-vs-recovered scatter at 10 repetitions, and panels B and C show recovery (correlation) and mean absolute error as a function of the number of repetitions per fact. The recovered offsets are stored as a named list-column in the fits, so `get_delta_alpha()` unpacks them into a long table keyed by fact before merging on the true values.

```{r}
# Extract true fact offsets
true_offsets <- sim_test[, .(true_delta_alpha = true_fact_offset[1]),
                         by = .(user_id, fact_id)]

# Extract recovered offsets from fits
get_delta_alpha <- function(fits_df) {
  fits_df[, {
    da <- delta_alpha_est[[1]]
    if (is.null(da) || length(da) == 0) {
      NULL
    } else {
      fact_ids <- as.integer(sub("^d_", "", names(da)))
      .(fact_id = fact_ids,
        delta_alpha_est = as.numeric(da))
    }
  }, by = .(user_id, n_repetitions)]
}

fact_recovery <- get_delta_alpha(fits_vary_reps) %>% filter(n_repetitions < 20)  
fact_recovery <- merge(fact_recovery, true_offsets, by = c("user_id", "fact_id"))

# (A) Scatter at 10 reps
fact_10 <- fact_recovery[n_repetitions == 10]
r_val <- cor(fact_10$true_delta_alpha, fact_10$delta_alpha_est, use = "complete.obs")

p_fact_scatter <- ggplot(fact_10, aes(x = true_delta_alpha, y = delta_alpha_est)) +
  geom_point(alpha = 0.7, color = pal_aaas()(1)) +
  geom_smooth(method = "lm", color = "black", se = TRUE) +
  annotate("text", x = -Inf, y = Inf,
           label = sprintf("r = %.2f", r_val),
           hjust = -0.1, vjust = 2, size = 4) +
  coord_cartesian(clip = "off") +
  labs(x = expression("True "*Delta*alpha),
       y = expression("Recovered "*Delta*alpha)) +
  plottheme

# (B) Correlation over reps
fact_corr <- fact_recovery[, .(corr = cor(true_delta_alpha, delta_alpha_est, use = "complete.obs")),
                           by = n_repetitions]

p_fact_corr <- ggplot(fact_corr, aes(x = factor(n_repetitions), y = corr, group = 1)) +
  geom_line(linewidth = 1, color = pal_aaas()(1)) +
  geom_point(size = 2, color = pal_aaas()(1)) +
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = "Repetitions per fact", y = "Recovery (correlation r)") +
  plottheme 

# (C) Error over reps
fact_error <- fact_recovery[, .(
  mean_error = mean(abs(true_delta_alpha - delta_alpha_est), na.rm = TRUE),
  se_error = sd(abs(true_delta_alpha - delta_alpha_est), na.rm = TRUE) / sqrt(.N)
), by = n_repetitions]

p_fact_error <- ggplot(fact_error, aes(x = factor(n_repetitions), y = mean_error, group = 1)) +
  geom_line(linewidth = 1, color = pal_aaas()(1)) +
  geom_point(size = 2, color = pal_aaas()(1)) +
  geom_errorbar(aes(ymin = mean_error - se_error, ymax = mean_error + se_error),
                width = 0.3, alpha = 0.6, color = pal_aaas()(1)) +
  labs(x = "Repetitions per fact", y = "Mean absolute error") +
  plottheme

(p_fact_scatter + p_fact_corr + p_fact_error) +
  plot_annotation(tag_levels = "A") &
  theme(
    plot.tag = element_text(face = "bold", size = 14)
  )

# Paper Figure 6: recovery of fact-level Speed of Forgetting offsets (delta-alpha)
ggsave(here::here("output", "fact_level_recovery.png"), width = 10, height = 3.5)
```

# Session info
```{r}
sessionInfo()
```
