---
title: "Example application: Modelling memory function in MCI and healthy ageing"
subtitle: "Idiographic Memory Modelling in ACT-R using Alternating Maximum Likelihood Estimation"
author: "Maarten van der Velde & Thomas Wilschut"
date: "Last updated: `r Sys.Date()`"
output:
  github_document:
    toc: yes
---

# Overview

In this notebook, we apply the AMLE method to an existing retrieval practice data set from [Hake et al. (2024)](https://www.medrxiv.org/content/10.1101/2024.03.15.24304345v1).
Participants completed multiple retrieval practice sessions over a period of about a year, resulting in repeated longitudinal measurements of memory performance.
The sample included participants clinically diagnosed with Mild Cognitive Impairment (MCI) and age-matched healthy control (HC) participants.
MCI is known to affect both memory and non-memory cognitive processes, which is visible in participants' performance on the task: Hake et al. found that participants' clinical status could be reliably determined from the Speed of Forgetting ($\phi$) parameter estimated from the data.
However, in that analysis, other parameters of the memory model were held constant.
Here, we fit the same data set using the AMLE method while allowing more parameters to vary, to see whether performance differences between individuals and groups are also explained through differences in other parameters (specifically: activation noise ($s$), retrieval threshold ($\tau$), latency factor ($F$) and non-retrieval time ($t_{er}$)).


# Setup

```{r load-packages}
library(here)
library(data.table)
library(purrr)
library(furrr)
library(tidyr)
library(jsonlite)
library(lmerTest)
library(psych)
library(ggplot2)
library(patchwork)
library(ggExtra)
library(GGally)
library(tidytext)

# Set up parallel processing
plan(multisession, workers = 8)

source(here("R", "sim-mle.R"))

set.seed(2026)

# Define colours
col_blue <- "#0571b0"
col_red <- "#ca0020"
col_green <- "#1b9e77"

# Define plotting theme
theme_set(theme_bw())

# Use cached version of model fit if available?
refit_model <- FALSE
```


# Load data

```{r load-data}
d <- fread(here("data", "processed", "hake2024.csv"))
```

# Inspect data

## Structure

Each row in the data corresponds to a single retrieval practice trial, recording the context in which it happened and the participant's performance:
```{r preview-data}
head(d)
```

## Session information

Participants completed multiple sessions, each of which lasted about 8 minutes.
The number of facts encountered, total number of trials, and performance varied across sessions depending on performance, which may be relevant when interpreting the fitted parameters later on.
```{r session-stats}
session_stats <- d[, .(
  trials = .N,
  errors = sum(correct == FALSE),
  accuracy = mean(correct),
  median_rt = median(rt),
  facts = uniqueN(fact_id),
  reps_per_fact = .N / uniqueN(fact_id),
  duration_min = ((max(start_time) + rt[which.max(start_time)]) - min(start_time))/60),
  by = .(user_id, clinical_status, session, session_id, lesson_id)]

stats_by_clinical_status <- session_stats[, .(
  sessions = .N,
  participants = uniqueN(user_id),
  mean_trials_per_session = mean(trials),
  mean_duration_min = mean(duration_min),
  mean_reps_per_fact = mean(reps_per_fact),
  mean_n_facts = mean(facts),
  mean_accuracy = mean(accuracy),
  mean_median_rt = mean(median_rt)
), by = clinical_status]

stats_combined <- session_stats[, .(
  clinical_status = "Combined",
  sessions = .N,
  participants = uniqueN(user_id),
  mean_trials_per_session = mean(trials),
  mean_duration_min = mean(duration_min),
  mean_reps_per_fact = mean(reps_per_fact),
  mean_n_facts = mean(facts),
  mean_accuracy = mean(accuracy),
  mean_median_rt = mean(median_rt)
)]

rbind(stats_by_clinical_status, stats_combined)
```

The pairwise plots below show how these session characteristics relate to each other, and how they differ between clinical groups.
On average, participants in the MCI group tended to complete fewer trials per session, had fewer error-free sessions and lower accuracy, had longer response times, and encountered fewer unique facts.
```{r session-stats-correlation, fig.height = 10, fig.width = 10}
ggpairs(session_stats,
        columns = c("trials", "errors", "accuracy", "median_rt", "facts", "reps_per_fact", "duration_min"),
        columnLabels = c("Trials", "Errors", "Accuracy", "Median RT (s)", "Unique facts", "Repetitions per fact", "Duration (min)"),
        mapping = aes(colour = clinical_status, fill = clinical_status, alpha = .25),
        upper = list(continuous = wrap("cor", method = "pearson")),
        progress = FALSE) +
  scale_color_manual(values = c(col_blue, col_red)) +
  scale_fill_manual(values = c(col_blue, col_red))
```

# Filter data

For model fitting, we'll apply the filtering rules used in Hake et al.'s analysis of this data set:

  - Exclude fact sequences containing a response time below 300 ms or above 15 s, which are likely to reflect accidental responding or disengagement.
  - Exclude fact sequences with fewer than 3 repetitions per fact, which contain insufficient information for reliable parameter estimation.

```{r filter-data}
# Filtering criteria
min_rt <- 0.3
max_rt <- 15
min_reps_per_fact <- 3

fact_sequences <- split(d, by = c("user_id", "session_id", "fact_id"))
fact_sequences_filtered <- fact_sequences[map_lgl(fact_sequences, ~ nrow(.x) >= min_reps_per_fact && min(.x$rt) >= min_rt && max(.x$rt) <= max_rt)]

d_filtered <- rbindlist(fact_sequences_filtered)
setorder(d_filtered, user_id, start_time)
```

Recalculate session statistics for the filtered set:
```{r session-stats-filtered, fig.height = 10, fig.width = 10}
session_stats_filtered <- d_filtered[, .(
  trials = .N,
  errors = sum(correct == FALSE),
  accuracy = mean(correct),
  median_rt = median(rt),
  facts = uniqueN(fact_id),
  reps_per_fact = .N / uniqueN(fact_id),
  duration_min = ((max(start_time) + rt[which.max(start_time)]) - min(start_time))/60),
  by = .(user_id, clinical_status, session, session_id, lesson_id)]

stats_by_clinical_status_filtered <- session_stats_filtered[, .(
  sessions = .N,
  participants = uniqueN(user_id),
  mean_trials_per_session = mean(trials),
  mean_duration_min = mean(duration_min),
  mean_reps_per_fact = mean(reps_per_fact),
  mean_n_facts = mean(facts),
  mean_accuracy = mean(accuracy),
  mean_median_rt = mean(median_rt)
), by = clinical_status]

stats_combined_filtered <- session_stats_filtered[, .(
  clinical_status = "Combined",
  sessions = .N,
  participants = uniqueN(user_id),
  mean_trials_per_session = mean(trials),
  mean_duration_min = mean(duration_min),
  mean_reps_per_fact = mean(reps_per_fact),
  mean_n_facts = mean(facts),
  mean_accuracy = mean(accuracy),
  mean_median_rt = mean(median_rt)
)]

rbind(stats_by_clinical_status_filtered, stats_combined_filtered)

ggpairs(session_stats,
        columns = c("trials", "errors", "accuracy", "median_rt", "facts", "reps_per_fact", "duration_min"),
        columnLabels = c("Trials", "Errors", "Accuracy", "Median RT (s)", "Unique facts", "Repetitions per fact", "Duration (min)"),
        mapping = aes(colour = clinical_status, fill = clinical_status, alpha = .25),
        upper = list(continuous = wrap("cor", method = "pearson")),
        progress = FALSE) +
  scale_color_manual(values = c(col_blue, col_red)) +
  scale_fill_manual(values = c(col_blue, col_red))
```

# Fit model

The data set is split into sessions, each of which is fitted separately.
```{r split-data}
d_sessions <- split(d_filtered, by = c("session_id"))
```

If desired, load previously fitted model results; otherwise, fit the model to each session and save the results for future use.
(Note: fitting takes a while!)
```{r fit-model}
delta_phi_file <- if (file.exists(here("data", "processed", "AMLE_delta_phi.csv"))) {
  here("data", "processed", "AMLE_delta_phi.csv")
} else {
  here("data", "processed", "AMLE_delta_alpha.csv")
}

if (refit_model == FALSE &&
    file.exists(here("data", "processed", "AMLE_fit.csv")) &&
    file.exists(delta_phi_file)) {
  fit_amle <- fread(here("data", "processed", "AMLE_fit.csv"))
  fit_delta_phi <- fread(delta_phi_file)
  # Backward compatibility: rename old column names if present
  if ("delta_alpha" %in% names(fit_delta_phi)) setnames(fit_delta_phi, "delta_alpha", "delta_phi")
  if ("alpha" %in% names(fit_amle)) setnames(fit_amle, "alpha", "phi")

} else {

  # Fit model
  fit_amle <- future_map(d_sessions, function (d_i) {
    
    fit_i_cpp <- fit_one_learner(d_i, full_iter = 100)[.N]
    
    # If the fit failed and returned NULL, replace with an empty data.table
    if(is.null(fit_i_cpp)) {
      fit_i_cpp <- data.table()
    }
    # Add participant and session information
    fit_i_cpp[, user_id := d_i$user_id[1]]
    fit_i_cpp[, clinical_status := d_i$clinical_status[1]]
    fit_i_cpp[, session := d_i$session[1]]
    fit_i_cpp[, session_id := d_i$session_id[1]]
    fit_i_cpp[, lesson_id := d_i$lesson_id[1]]
    fit_i_cpp[, session_trials := nrow(d_i)]
    fit_i_cpp[, session_errors := sum(d_i$correct == FALSE)]
    fit_i_cpp[, session_accuracy := mean(d_i$correct)]
    
    return(fit_i_cpp)
  }, .progress = interactive())
  
  fit_amle <- rbindlist(fit_amle, fill = TRUE)
  
  # Extract delta_phi estimates
  fit_delta_phi <- fit_amle[, .(delta_phi = delta_phi_est), by = .(user_id, clinical_status, session, session_id, lesson_id, session_trials, session_errors, session_accuracy)]
  fit_delta_phi[, delta_phi := lapply(delta_phi, function (delta_phi) {
    as.data.table(delta_phi, keep.rownames = "fact_id")
  })]
  fit_delta_phi <- unnest(fit_delta_phi, delta_phi) |> as.data.table()
  
  # Save results
  fwrite(fit_amle, here("data", "processed", "AMLE_fit.csv"))
  fwrite(fit_delta_phi, here("data", "processed", "AMLE_delta_phi.csv"))
  
}
```

# Evaluate model fit

## How well does the model fit capture the data?

How well does the fitted model capture the distribution of response times?
Here, we use the fitted parameters to predict trial-level response times.
```{r predict-rt-from-fit}
fitted_session_params <- fit_amle[, .(session_id, phi, tau, s, lf, ter)]
fitted_session_params <- fit_delta_phi[fitted_session_params, on = .(session_id)]
fitted_session_params[, fact_sof := phi + delta_phi]
fitted_session_params[, fact_id := gsub("d_", "", fact_id)]

d_preds <- future_map(d_sessions, function (d_session) {
  
  session_params <- fitted_session_params[session_id == d_session$session_id[1]]
  
  # If session params is empty or parameters are NA, return NULL
  if (nrow(session_params) == 0 || any(is.na(session_params[, .(delta_phi, phi, tau, s, lf, ter)]))) {
    return (NULL)
  }
  
  fitted_ter <- session_params[1, ter]
  fitted_lf <- session_params[1, lf]
  fitted_tau <- session_params[1, tau]
  
  pred_rt_error <- fitted_ter + fitted_lf * exp(-fitted_tau)
  
  setorder(d_session, start_time)
  
  model_predictions <- map(seq_len(nrow(d_session)), function (trial) {
    current_trial <- d_session[trial]
    previous_trials_for_fact <- d_session[fact_id == current_trial$fact_id & start_time < current_trial$start_time]
    
    # If there are no previous trials for this fact, the activation is -Inf
    if (nrow(previous_trials_for_fact) == 0) {
      activation <- -Inf
    } else {
      activation <- calculate_activation_memory(t = current_trial$start_time,
                                                traces = previous_trials_for_fact$start_time, 
                                                sof = session_params[fact_id == current_trial$fact_id, fact_sof])
    }
    pred_rt_correct <- fitted_ter + fitted_lf * exp(-activation)
    
    return (list(
      activation = activation,
      pred_rt_correct = pred_rt_correct,
      pred_rt_error = pred_rt_error))
    
  }) |> rbindlist()
  
  return (cbind(d_session, model_predictions))
  
}) |> rbindlist()
```

### Decile plots of predicted and observed RT

The decile plots below compare each participant's response time distribution (on correct responses) as predicted from fitted model parameters to their observed response time distribution.
Agreement is quantified per participant by a Pearson correlation, shown in the top of each plot.
```{r decile-plots, fig.height = 10, fig.width = 10}

# Create simplified user IDs that are sorted by clinical status
user_ids <- unique(d_preds[, .(user_id, clinical_status)])
setorder(user_ids, clinical_status)
user_ids[, user_id_simple := sprintf("%02d", .I)]
d_preds <- user_ids[d_preds, on = .(user_id, clinical_status)]

# Only evaluate RT predictions on correct test trials
d_preds_correct <- d_preds[repetition != 1 & correct == TRUE]

# Test the correlation between predicted and observed RT per participant
d_preds_correct_by_user <- split(d_preds_correct, by = "user_id_simple")

cor_rt <- map(d_preds_correct_by_user, function (x) {
  cor_rt <- cor.test(x$pred_rt_correct, x$rt)
  return (list(
    user_id_simple = x$user_id_simple[1],
    clinical_status = x$clinical_status[1],
    r = cor_rt$estimate,
    p = cor_rt$p.value,
    n = nrow(x)
  ))
}) |> rbindlist()
cor_rt[, cor_label := paste0("r(", n, ") = ", sprintf("%.2f", r), ifelse(p < .001, "***", ifelse(p < .01, "**", ifelse(p < .05, "*", ""))))]

# Also compute decile points to plot on top (i.e., 10th decile of predicted RTs and 10th decile of observed RTs)
deciles <- d_preds_correct[, .(
  pred_rt_correct = quantile(pred_rt_correct, probs = seq(.1, .9, by = .1)),
  rt = quantile(rt, probs = seq(.1, .9, by = .1))
), by = .(user_id_simple, clinical_status)]

# Create plot
p_decile_rt <- ggplot(d_preds_correct, aes(x = pred_rt_correct, y = rt, colour = clinical_status)) +
  facet_wrap(~ user_id_simple, ncol = 9, labeller = labeller(user_id_simple = function(x) paste("Participant", x))) +
  geom_point(alpha = .01, size = .5, colour = "grey30") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  geom_line(data = deciles, alpha = .5) +
  geom_point(data = deciles, alpha = .8) +
  geom_label(data = cor_rt, aes(x = 7.5, y = 15, label = cor_label), hjust = .5, vjust = 1, size = 2.5, fill = "white", show.legend = FALSE) +
  labs(x = "Predicted RT (s)", y = "Observed RT (s)", colour = "Clinical status") +
  scale_colour_manual(values = c("HC" = col_blue, "MCI" = col_red)) +
  coord_equal(ratio = 1, xlim = c(0, 15), ylim = c(0, 15)) +
  guides(colour = guide_legend(override.aes = list(alpha = 1))) +
  theme(legend.position = c(.8375, .075), legend.direction = "horizontal")

# Paper Figure 8B: per-participant decile plots of predicted vs observed RT
ggsave(here("output", "predicted_vs_observed_rt.png"), plot = p_decile_rt, width = 10, height = 10)
p_decile_rt
```

### Correlation between predicted and observed RT

The correlation coefficients between predicted and observed RTs (on correct test trials) are distributed as follows:
```{r correlation-rt, fig.height = 4, fig.width = 10}
cor_rt_dist <- copy(cor_rt)
cor_rt_dist[, clinical_status := factor(clinical_status, levels = c("MCI", "HC"))]

p_rt_corr <- ggplot(cor_rt_dist, aes(x = r, fill = clinical_status, y = clinical_status)) +
  # geom_violin(alpha = .5) +
  # ggdist::stat_halfeye(adjust = .5, justification = -.5, alpha = .5, scale = .5, .width = 0, point_colour = NA) +
  geom_boxplot(outlier.shape = NA, width = .2, alpha = .5) +
  geom_jitter(aes(colour = clinical_status), width = 0, height = .2, alpha = .25) +
  labs(x = "Correlation between predicted and observed RT", y = "Clinical status") +
  scale_x_continuous(limits = c(0, 1)) +
  scale_fill_manual(values = c("HC" = col_blue, "MCI" = col_red)) +
  scale_colour_manual(values = c("HC" = col_blue, "MCI" = col_red)) +
  guides(fill = "none", colour = "none")

# Paper Figure 8A: distribution of per-participant predicted-vs-observed RT correlations
ggsave(here("output", "predicted_vs_observed_rt_correlation_dist.png"), plot = p_rt_corr, width = 10, height = 4)
p_rt_corr
```

Combined version of these RT fit plots:
```{r rt-fit-plots, fig.height = 10, fig.width = 10}
p_rt_corr + p_decile_rt  +
  plot_layout(ncol = 1, heights = c(1, 8)) +
  plot_annotation(tag_levels = "A") &
  theme(plot.tag = element_text(face = "bold", size = 14))

# Paper Figure 8: RT model fit, panels A (correlation distribution) + B (decile plots) combined
ggsave(here("output", "rt_fit_plots.png"), width = 10, height = 10)
```

### Visual comparison of full RT distributions

We'll also compare the full RT distributions, including incorrect responses.
For incorrect RTs (predicted and observed) we flip the sign so that they appear on the left side of the plot.
```{r rt-distribution-full, fig.height = 10, fig.width = 10}
d_preds_both <- d_preds[repetition != 1]
d_preds_both[, pred_rt := ifelse(correct == TRUE, pred_rt_correct, pred_rt_error)]

# Flip the sign of error RTs
d_preds_both[, pred_rt := ifelse(correct == TRUE, pred_rt, -pred_rt)]
d_preds_both[, rt := ifelse(correct == TRUE, rt, -rt)]

ggplot(d_preds_both, aes(x = rt, fill = clinical_status)) +
  facet_wrap(~ user_id_simple, ncol = 9, labeller = labeller(user_id_simple = function(x) paste("Participant", x))) +
  geom_density(aes(alpha = "Observed")) +
  geom_density(aes(x = pred_rt, alpha = "Predicted")) +
  geom_vline(xintercept = 0, linetype = "dotted") +
  scale_x_continuous(limits = c(-15, 15)) +
  labs(x = "RT (s)", y = "Density", fill = "Clinical status") +
  scale_fill_manual(values = c("HC" = col_blue, "MCI" = col_red)) +
  scale_alpha_manual(name = "Distribution", values = c("Observed" = .15, "Predicted" = .75), labels = c("Observed", "Predicted")) +
  theme(legend.position = c(.8375, .075), legend.direction = "horizontal") +
  guides(fill = guide_legend(override.aes = list(alpha = 1)),
         alpha = guide_legend(override.aes = list(fill = "grey20")))

# Paper Figure 7: predicted vs observed RT density per participant (negative RTs = incorrect)
ggsave(here("output", "predicted_vs_observed_rt_density.png"), width = 10, height = 8.5)
```


## Estimated parameters

### Parameter estimates per session
```{r parameter-estimates}
setorder(fit_amle, clinical_status, user_id, session, lesson_id)
fit_amle[, session_aligned := 1:.N, by = .(user_id)]
fit_amle_long <- melt(fit_amle, measure.vars = c("phi", "tau", "s", "lf", "ter"))
fit_amle_long <- user_ids[fit_amle_long, on = .(user_id, clinical_status)]
```

The following plot shows each session's parameter estimates, with a horizontal line indicating the average parameter value across sessions for each participant.
```{r parameter-estimates-over-time, fig.height = 60, fig.width = 10}
fit_amle_long_avg <- fit_amle_long[, .(value_mean = mean(value, na.rm = TRUE)), by = .(user_id_simple, clinical_status, variable)]

ggplot(fit_amle_long, aes(x = session_aligned, y = value, colour = clinical_status, group = user_id_simple)) +
  facet_wrap(user_id_simple ~ variable, ncol = 5, scale = "free_y", labeller = labeller(user_id_simple = function(x) paste("Participant", x))) +
  geom_point(alpha = .2) +
  geom_hline(data = fit_amle_long_avg, aes(yintercept = value_mean, colour = clinical_status), lwd = 1) +
  labs(x = "Session number (aligned)", y = "Parameter value", colour = "Clinical\nstatus") +
  scale_colour_manual(values = c("HC" = col_blue, "MCI" = col_red)) +
  scale_x_continuous(breaks = seq(0, 60, by = 30)) +
  theme(legend.position = "bottom")
```

### Relationship to session properties

How does the number of trials, number of unique facts, number of errors, accuracy or RT in a session relate to the parameter estimates?
```{r parameter-estimates-session-stats, fig.height = 6, fig.width = 10}
# Add session stats
fit_amle_long <- session_stats[fit_amle_long, on = .(user_id, clinical_status, session, session_id, lesson_id)]

ggplot(fit_amle_long, aes(x = session_trials, y = value)) +
  facet_wrap(clinical_status ~ variable, scales = "free_y", ncol = 5) +
  geom_point(alpha = .2) +
  geom_smooth(method = "gam") +
  labs(x = "Number of trials in session", y = "Parameter value", title = "Session trials")

ggplot(fit_amle_long, aes(x = facts, y = value)) +
  facet_wrap(clinical_status ~ variable, scales = "free_y", ncol = 5) +
  geom_point(alpha = .2) +
  geom_smooth(method = "gam") +
  labs(x = "Number of unique facts in session", y = "Parameter value", title = "Unique facts")

ggplot(fit_amle_long, aes(x = session_errors, y = value)) +
  facet_wrap(clinical_status ~ variable, scales = "free_y", ncol = 5) +
  geom_point(alpha = .2) +
  geom_smooth(method = "gam") +
  labs(x = "Number of errors in session", y = "Parameter value", title = "Session errors") 

ggplot(fit_amle_long, aes(x = session_accuracy, y = value)) +
  facet_wrap(clinical_status ~ variable, scales = "free_y", ncol = 5) +
  geom_point(alpha = .2) +
  geom_smooth(method = "gam") +
  labs(x = "Session accuracy", y = "Parameter value", title = "Session accuracy")

ggplot(fit_amle_long, aes(x = median_rt, y = value)) +
  facet_wrap(clinical_status ~ variable, scales = "free_y", ncol = 5) +
  geom_point(alpha = .2) +
  geom_smooth(method = "gam") +
  labs(x = "Median RT in session (s)", y = "Parameter value", title = "Session median RT")
```

### Stability over time

Participants completed multiple sessions over time, and parameters are estimated independently for each session.
How consistent are parameter estimates within each participant?
As we would expect, there is variation in estimated parameter values from session to session.
This variation can come from three sources:

  1. **Materials**: Part of this variation will be due to participants practicing different materials in each session, meaning that some sessions will end up being a bit more difficult than others.
  2. **Participants**: Part of this variation will also be within-participant variability, due to fluctuations in cognitive performance over time.
  3. **Noise**: Finally, part of this variation will be noise, e.g. due to measurement error or sources of variance that are otherwise not accounted for.
  
To find out how consistent estimates are within participants (more precisely: what proportion of variance can be attributed to stable differences between participants), we can compute intra-class correlations (ICC).
To factor out variance due to fact/lesson differences and to account for missing data, we'll calculate ICC from a mixed-effects regression model with random intercepts for participants and lessons.

$$ICC_{participant} = \frac{\sigma^2_{participant}}{\sigma^2_{participant} + \sigma^2_{residual}}$$

NB: this assumes that parameters are stationary (i.e., don't systematically change over time).
If that isn't the case, this method underestimates the true consistency.

The result is shown per parameter under ICC3 in the tables below.
```{r}
for (param in fit_amle_long[, unique(variable)]) {
  icc_data_param <- dcast(fit_amle_long[variable == param], user_id_simple ~ lesson_id, value.var = "value")
  icc_param <- ICC(icc_data_param[, -1, with = FALSE], lmer = TRUE)
  print(paste("Parameter:", param))
  print(icc_param$results)
}
```

### Average estimated parameters

Aggregate estimates within participants to get an overall average.
How do these differ within and between groups?

```{r average-parameters}
fit_amle_long_avg <- fit_amle_long[, .(value_mean = mean(value, na.rm = TRUE)), by = .(user_id_simple, clinical_status, variable)]
```
Test: Is there a significant difference in parameters between participants of different clinical status?

Yes: phi, lf, ter.
No: tau, s.
```{r parameter-group-diff-lmer}
for (param in fit_amle_long[, unique(variable)]) {
  model <- lmer(value ~ clinical_status + (1 | user_id) + (1 | lesson_id), data = fit_amle_long[variable == param])
  summary_model <- summary(model)
  print(paste("Parameter:", param))
  print(summary_model)
}
```

### Correlation

Are parameters correlated?
```{r}
fit_amle_avg <- dcast(fit_amle_long_avg, user_id_simple + clinical_status ~ variable, value.var = "value_mean")

fit_amle_avg_plot <- fit_amle_avg[, .(
  phi = mean(phi, na.rm = TRUE),
  tau = mean(tau, na.rm = TRUE),
  s = mean(s, na.rm = TRUE),
  lf = mean(lf, na.rm = TRUE),
  ter = mean(ter, na.rm = TRUE),
  n_sessions = sum(!is.na(phi))),
  by = .(user_id = user_id_simple, clinical_status)]

param_cols <- setdiff(names(fit_amle_avg_plot), c("user_id", "clinical_status", "n_sessions"))

fit_amle_avg_plot[, (param_cols) := lapply(.SD, scale), .SDcols = param_cols]

fit_amle_avg_plot <- melt(
  fit_amle_avg_plot,
  id.vars = c("user_id", "clinical_status"),
  measure.vars = param_cols,
  variable.name = "parameter",
  value.name = "z"
)


# Custom ticks per parameter (raw values)
ticks_list <- list(
  phi = c(0.1, 0.3, 0.5, 0.7, 0.9),
  tau = c(-4, -3, -2, -1),
  s   = c(0.1, 0.2, 0.3, 0.4),
  lf  = c(0, 1, 2, 3),
  ter = c(.5, 1, 1.5, 2, 2.5)
)

# Map ticks to a long data.table for annotation
axis_anno <- rbindlist(
  lapply(names(ticks_list), function(var) {
    data.table(
      variable = var,
      q = ticks_list[[var]]
    )
  })
)

# Compute z-scores for plotting
param_stats <- fit_amle_long_avg[, .(mean = mean(value_mean), sd = sd(value_mean)), by = .(variable)]
axis_anno <- merge(axis_anno, param_stats, by = "variable")
axis_anno[, z := (q - mean) / sd]

# Map variables to x-axis positions
axis_anno[, x := as.numeric(factor(variable, levels = param_cols))]

# Compute group means
group_means <- fit_amle_avg_plot[
  , .(z = mean(z)), 
  by = .(clinical_status, parameter)
]

# Map x-axis positions
group_means[, x := as.numeric(factor(parameter, levels = param_cols))]

# Add jitter
fit_amle_avg_plot[, x_num := as.numeric(factor(fit_amle_avg_plot$parameter, levels = param_cols))]
fit_amle_avg_plot[, x_jitter := runif(1, -0.05, 0.05), by = user_id]
fit_amle_avg_plot[, x := x_num + x_jitter]



# Numeric x positions for parameters
param_x <- seq_along(param_cols)

# Max y value to position labels above the plot
y_max <- max(c(fit_amle_avg_plot$z, group_means$z))

# Create a dataframe for top labels
param_titles <- list( phi = "Speed of Forgetting", tau = "Retrieval threshold", s = "Activation noise", lf = "Latency factor", ter = "Non-retrieval time" )
param_symbols <- list( phi = "phi", tau = "tau", s = "s", lf = "F", ter = "t[er]" )

param_label_df <- data.frame(
  x = param_x,
  title = unlist(param_titles),
  symbol = unlist(param_symbols)
)

# Plot

p_param_estimates <- ggplot() +

  # Vertical axes
  geom_vline(xintercept = 1:length(param_cols), color = "grey80") +

  # Fake tick marks
  geom_segment(
    data = axis_anno,
    aes(x = x - 0.05, xend = x + 0.05,y = z,yend = z),
    inherit.aes = FALSE,
    color = "grey50"
  ) +

  # Individual participant lines (behind)
  geom_line(
    data = fit_amle_avg_plot,
    aes(x = x, y = z, group = user_id, color = factor(clinical_status)),
    alpha = 0.25
  ) +

  # Individual points
  geom_point(
    data = fit_amle_avg_plot,
    aes(x = x, y = z, group = user_id, color = factor(clinical_status)),
    alpha = 0.25,
    size = 1
  ) +

  # Group mean lines (on top)
  geom_line(
    data = group_means,
    aes(x = x, y = z, color = factor(clinical_status)),
    linewidth = 1
  ) +

  # Group mean points
  geom_point(
    data = group_means,
    aes(x = x, y = z, color = factor(clinical_status)),
    size = 2
  ) +
  
  # Fake tick labels on the right
  geom_label(
    data = axis_anno,
    aes(x = x + 0.1, y = z, label = q),
    size = 3,
    hjust = 0,
    alpha = .5,
    label.size = 0,
    label.padding = unit(c(0,0,0,0), "lines"),
    inherit.aes = FALSE
  ) +
  
  geom_text(
    data = axis_anno,
    aes(x = x + 0.1, y = z, label = q),
    size = 3,
    hjust = 0,
    inherit.aes = FALSE
  ) +

  # X-axis: parameter names along top
  geom_label(
    data = param_label_df,
    aes(x = x, y = 6, label = paste0("atop(bold('", title, "'), ", symbol, ")")),
    parse = TRUE,
    size = 3.5,
    vjust = 1,
    label.size = 0,
    label.padding = unit(c(.5, 0, .5, 0), "lines")
  ) +
  
  scale_x_continuous(
    expand = expansion(mult = .1),
    breaks = param_x
  ) +

  theme_minimal() +
  # Remove y-axis
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    strip.text.x = element_text(size = 10)) +

  # Labels
  labs(
    x = "Parameter",
    color = "Clinical status"
  ) +
  scale_colour_manual(values = c("HC" = col_blue, "MCI" = col_red))

# Not in paper: per-participant parameter estimates (exploratory)
ggsave(here("output", "parameter_estimates_by_participant.png"), width = 10, height = 5)
```

Plot parameters separately:
```{r parameter-estimates-boxplot, fig.height = 6, fig.width = 10}
p_parameter_estimates <- ggplot(fit_amle_long_avg, aes(x = variable, y = value_mean, fill = clinical_status)) +
  facet_wrap(~ variable, scales = "free", ncol = 5, labeller = labeller(variable = c(phi = "Speed of Forgetting", tau = "Retrieval threshold", s = "Activation noise", lf = "Latency factor", ter = "Non-retrieval time"))) +
  geom_boxplot(outlier.shape = NA, alpha = .5) +
  geom_point(position = position_jitterdodge(jitter.height = 0), alpha = .5, aes(colour = clinical_status)) +
  labs(x = "Parameter", y = "Estimated value", colour = "Clinical status", fill = "Clinical status") +
  scale_colour_manual(values = c("HC" = col_blue, "MCI" = col_red)) +
  scale_fill_manual(values = c("HC" = col_blue, "MCI" = col_red)) +
  scale_x_discrete(labels = c(phi = expression(phi), tau = expression(tau), s = expression(s), lf = expression(F), ter = expression(t[er]))) +
  theme(legend.position = "bottom")

# Paper Figure 9A: estimated parameters by clinical status (boxplots)
ggsave(here("output", "parameter_estimates_boxplot.png"), plot = p_parameter_estimates, width = 10, height = 6)
p_parameter_estimates
```
Pairwise correlations between parameters:
```{r parameter-estimates-correlations, fig.height = 10, fig.width = 10}
p_pairwise_corrs <- ggpairs(
  fit_amle_avg,
  columns = param_cols,
  columnLabels = unlist(param_titles),
  mapping = aes(color = clinical_status, fill = clinical_status, alpha = 0.5),
  legend = c(1,1),
  upper = list(continuous = wrap("cor", digits = 2))
) +
  scale_color_manual(values = c("HC" = col_blue, "MCI" = col_red)) +
  scale_fill_manual(values = c("HC" = col_blue, "MCI" = col_red)) +
  guides(alpha = "none",
         fill = guide_legend(override.aes = list(alpha = .5))) +
  labs(color = "Clinical status", fill = "Clinical status") +
    theme(legend.position = "bottom")

p_pairwise_corrs
```

Combine into one plot:
```{r parameter-estimates-combined, fig.height = 12, fig.width = 10}
p_combined <- p_parameter_estimates / wrap_elements(ggmatrix_gtable(p_pairwise_corrs)) +
  plot_layout(ncol = 1, heights = c(1, 4)) +
  plot_annotation(tag_levels = "A") &
  theme(plot.tag = element_text(face = "bold", size = 14))

# Paper Figure 9: parameter estimates by clinical status (A) + pairwise correlations (B)
ggsave(here("output", "parameter_estimates_correlations.png"), plot = p_combined, width = 10, height = 12)

p_combined
```

## Fact-level offsets

The relative difficulty of facts was estimated through offsets to a participant's $\phi$ parameter: $\Delta\phi$.
In the fitting process, these offsets were continually re-centered on zero, meaning that the average $\Delta\phi$ across facts for each participant is (approximately) zero, which makes it easier to compare them across participants.

The plot below shows the mean $\Delta\phi$ for each fact (+/- 1 SD), averaged across participants who encountered it, and organised by lesson.
```{r delta-phi-by-lesson, fig.height = 10, fig.width = 10}
fit_delta_phi_avg <- fit_delta_phi[, .(.N, delta_phi_mean = mean(delta_phi), delta_phi_sd = sd(delta_phi)), by = .(lesson_id, fact_id)]

ggplot(fit_delta_phi_avg, aes(x = reorder_within(fact_id, delta_phi_mean, lesson_id, min), y = delta_phi_mean)) +
  geom_errorbar(aes(ymin = delta_phi_mean - delta_phi_sd, ymax = delta_phi_mean + delta_phi_sd), width = 0, alpha = .25) +
  geom_point(aes(alpha = N)) +
  facet_wrap(~ lesson_id, scales = "free_x", ncol = 8) +
  scale_x_reordered() +
  geom_hline(yintercept = 0, colour = col_green) +
  labs(x = "Fact", y = expression(Delta~phi)) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major.x = element_blank())
```

Some facts/lessons were only encountered by a small number of participants:
```{r num-participants-per-fact, fig.height = 4, fig.width = 10}
ggplot(fit_delta_phi_avg, aes(x = N)) +
  geom_histogram(binwidth = 1) +
  labs(x = "Number of participants who encountered the fact", y = "Count")
```
To quantify the agreement in relative difficulty across participants, we can compute the intra-class correlation (ICC) of $\Delta\phi$ estimates across facts.
However, the fact that each participant's delta phi estimates center around zero means that a mixed-effects model with random intercepts for participants cannot be fitted to the data, as the participant-level variance is effectively zero.
We'll calculate the net phi value for each fact by adding the participant's overall phi to the fact-specific delta phi:
```{r delta-phi-by-lesson-2, fig.height = 10, fig.width = 10}
user_phi <- fit_amle_long[variable == "phi", .(user_id, session_id, user_phi = value)]
fit_delta_phi <- user_phi[fit_delta_phi, on = .(user_id, session_id)]
fit_delta_phi[, fact_phi := user_phi + delta_phi]

ggplot(fit_delta_phi, aes(x = reorder_within(fact_id, fact_phi, lesson_id, mean), y = fact_phi)) +
  geom_line(aes(group = user_id), alpha = .2) +
  facet_wrap(~ lesson_id, scales = "free_x", ncol = 8) +
  scale_x_reordered() +
  labs(x = "Fact", y = expression(Delta~phi)) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid.major.x = element_blank())
```

Then, compute ICC from a mixed-effects regression model with random intercepts for participants and lessons:
```{r delta-phi-icc}
delta_phi_wide <- dcast(fit_delta_phi, fact_id ~ user_id, value.var = "fact_phi")
icc_delta_phi <- ICC(delta_phi_wide[, -1, with = FALSE], lmer = TRUE)
print(icc_delta_phi$results)
```


