---
title: "Example application: parameter contributions to MCI vs HC classification"
subtitle: "Idiographic Memory Modelling in ACT-R using Alternating Maximum Likelihood Estimation"
author: "Thomas Wilschut"
date: "Last updated: `r Sys.Date()`"
output:
  github_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width  = 8,
  fig.height = 5,
  dpi = 150
)

library(data.table)
library(here)
library(tidyverse)
library(furrr)
library(future)
library(pROC)
library(ggplot2)
library(ggsci)
library(patchwork)
```

# Setup and helpers

The classification analysis below reports performance as a bootstrap-corrected AUC, to guard against the optimism of evaluating a model on the same data it was fit to. We first set up the parameter lookups and the bootstrap helper used throughout.

## Helpers

A few display lookups (full parameter names and a consistent colour palette) that are reused across the plots below.

```{r}
# ── Shared constants ──────────────────────────────────────────────────────────
PARAM_COL <- c(phi = "phi", tau = "tau", s = "s", F = "lf", ter = "ter")

param_display <- c(
  phi = "Speed of Forgetting",
  tau = "Retrieval threshold",
  s   = "Activation noise",
  F   = "Latency factor",
  ter = "Non-retrieval time"
)

param_cols <- setNames(
  pal_aaas()(5),
  c("Speed of Forgetting", "Retrieval threshold", "Activation noise",
    "Latency factor", "Non-retrieval time")
)
```


The idea is to first compute the apparent AUC (the model evaluated on the same data it was fit to, which is optimistically biased), then estimate that optimism by repeatedly refitting on bootstrap resamples and re-evaluating on the original data; the corrected AUC subtracts the mean optimism. `bootstrap_auc_corrected()` implements this and is the function used inside the greedy selection below.

```{r}
# ── Bootstrap-corrected AUC (general, used in greedy selection) ───────────────
bootstrap_auc_corrected <- function(outcome, predictors_df,
                                    n_boot = 500, seed = 42) {
  set.seed(seed)
  data   <- data.frame(outcome = outcome, predictors_df)
  n      <- nrow(data)
  params <- names(predictors_df)

  fmla <- if (length(params) == 1) NULL
          else as.formula(paste("outcome ~", paste(params, collapse = " + ")))

  if (is.null(fmla)) {
    roc_app <- roc(data$outcome, data[[params]], direction = "auto", quiet = TRUE)
    p_app   <- data[[params]]
  } else {
    fit_full <- glm(fmla, data = data, family = binomial())
    p_app    <- predict(fit_full, type = "response")
    roc_app  <- roc(data$outcome, p_app, direction = "auto", quiet = TRUE)
  }
  auc_app <- as.numeric(auc(roc_app))

  boot_aucs <- replicate(n_boot, {
    idx    <- sample(n, n, replace = TRUE)
    b_data <- data[idx, ]
    if (length(unique(b_data$outcome)) < 2) return(NA_real_)
    if (is.null(fmla)) {
      p_orig <- data[[params]]
    } else {
      fit_b <- tryCatch(glm(fmla, data = b_data, family = binomial()),
                        error = function(e) NULL)
      if (is.null(fit_b)) return(NA_real_)
      p_orig <- predict(fit_b, newdata = data, type = "response")
    }
    tryCatch(
      as.numeric(auc(roc(data$outcome, p_orig, direction = "auto", quiet = TRUE))),
      error = function(e) NA_real_
    )
  })

  optimism      <- auc_app - mean(boot_aucs, na.rm = TRUE)
  auc_corrected <- auc_app - optimism

  list(
    auc_apparent  = auc_app,
    auc_corrected = auc_corrected,
    auc_lo        = quantile(boot_aucs, 0.025, na.rm = TRUE),
    auc_hi        = quantile(boot_aucs, 0.975, na.rm = TRUE)
  )
}
```


# Greedy forward selection with refitting

To see which parameters carry the diagnostic signal, we perform a greedy forward
selection: starting from no free parameters, at each step we add the parameter
that most improves classification. Crucially, for each candidate parameter set
the data are refit with the non-selected parameters held fixed at their
population defaults, so that each model genuinely uses only the information in
its selected parameters.

## Data preparation

For the refitting analysis we reload the raw per-trial data and apply the same filtering rules as Hake et al. (drop fact sequences with fewer than three encounters, or with response times below 300 ms or above 15 s), and code the outcome as MCI = 1 / HC = 0.

```{r}
future::plan(future::multisession, workers = future::availableCores() - 1)

source(here::here("R", "sim-mle.R"))
d <- fread(here::here("data", "processed", "hake2024.csv"))

min_rt             <- 0.3
max_rt             <- 15
min_reps_per_fact  <- 3

fact_sequences <- split(d, by = c("user_id", "session_id", "fact_id"))
fact_sequences_filtered <- fact_sequences[
  map_lgl(fact_sequences, ~ nrow(.x) >= min_reps_per_fact &&
            min(.x$rt) >= min_rt && max(.x$rt) <= max_rt)
]

d_filtered <- rbindlist(fact_sequences_filtered)
setorder(d_filtered, user_id, start_time)

clinical_df <- unique(d_filtered[, .(user_id, clinical_status)])[
  , outcome := as.integer(clinical_status == "MCI")
]
```


## Greedy selection: person-average criterion

`greedy_select_auc()` performs the forward selection: starting from no free parameters, at each step it refits every session with each remaining candidate added to the already-selected set, aggregates the session fits to person-level means, and keeps the candidate that yields the highest bootstrap-corrected AUC. It returns both the per-step summary and the underlying fits (the latter are reused below to evaluate the single-session criterion without refitting).

```{r}
# ── Aggregate session-level fits to person-level means ────────────────────────
person_means <- function(fits_dt, free_params, clinical_df) {
  cols <- PARAM_COL[free_params]
  cols <- cols[cols %in% names(fits_dt)]
  agg  <- fits_dt[, lapply(.SD, mean, na.rm = TRUE),
                  by = .(user_id), .SDcols = cols]
  merge(agg, clinical_df, by = "user_id")
}

# ── Greedy forward selection (person-average AUC criterion) ───────────────────
greedy_select_auc <- function(d_filtered,
                              all_params = c("phi", "tau", "s", "F", "ter"),
                              n_boot     = 500,
                              full_iter  = 100) {

  clinical_df <- unique(d_filtered[, .(user_id, clinical_status)])[
    , outcome := as.integer(clinical_status == "MCI")
  ]
  d_sessions <- split(d_filtered, by = "session_id")

  fit_sessions <- function(free_params) {
    message("  Fitting ", length(d_sessions), " sessions with free: {",
            paste(free_params, collapse = ", "), "} ...")
    results <- future_map(d_sessions, function(d_i) {
      fit_i <- fit_one_learner(d_i, free_params = free_params,
                               full_iter = full_iter)[.N]
      if (is.null(fit_i) || nrow(fit_i) == 0) return(data.table())
      fit_i[, `:=`(user_id         = d_i$user_id[1],
                   clinical_status = d_i$clinical_status[1],
                   session_id      = d_i$session_id[1])]
      fit_i
    }, .options = furrr_options(seed = TRUE), .progress = interactive())
    results <- results[!vapply(results, function(x) is.null(x) || nrow(x) == 0,
                               logical(1))]
    rbindlist(results, fill = TRUE)
  }

  selected  <- character(0)
  remaining <- all_params
  log_rows  <- list()
  all_fits  <- list()

  for (step in seq_along(all_params)) {
    message("\n=== Step ", step, ": testing ", length(remaining), " candidate(s) ===")
    step_rows <- list()

    for (candidate in remaining) {
      trial_params <- c(selected, candidate)
      message("  Candidate: {", paste(trial_params, collapse = ", "), "}")

      fits_dt <- fit_sessions(trial_params)
      if (nrow(fits_dt) == 0) { message("  WARNING: no fits returned"); next }

      pm        <- person_means(fits_dt, trial_params, clinical_df)
      pred_cols <- PARAM_COL[trial_params]
      pred_cols <- pred_cols[pred_cols %in% names(pm)]
      auc_res   <- bootstrap_auc_corrected(pm$outcome, pm[, ..pred_cols],
                                           n_boot = n_boot)

      message(sprintf("    AUC corrected = %.3f [%.3f-%.3f]",
                      auc_res$auc_corrected, auc_res$auc_lo, auc_res$auc_hi))

      step_rows[[candidate]] <- data.table(
        step = step, param_added = candidate, free_params = list(trial_params),
        auc_apparent = auc_res$auc_apparent, auc_corrected = auc_res$auc_corrected,
        auc_lo = auc_res$auc_lo, auc_hi = auc_res$auc_hi, selected = FALSE
      )

      key      <- paste(step, candidate, sep = "_")
      fit_copy <- copy(fits_dt)
      fit_copy[, `:=`(step = step, param_added = candidate,
                      free_params_str = paste(trial_params, collapse = ","))]
      all_fits[[key]] <- fit_copy
    }

    step_dt  <- rbindlist(step_rows)
    best_idx <- which.max(step_dt$auc_corrected)
    best_par <- step_dt$param_added[best_idx]
    step_dt[best_idx, selected := TRUE]

    message(sprintf("  >>> Selected: %s  (AUC = %.3f)",
                    best_par, step_dt$auc_corrected[best_idx]))

    log_rows[[step]] <- step_dt
    selected  <- c(selected, best_par)
    remaining <- setdiff(remaining, best_par)
  }

  list(summary = rbindlist(log_rows),
       learner_fits = rbindlist(all_fits, fill = TRUE))
}
```


```{r, eval=FALSE}
# Expensive (refits every session for every candidate); cached and reloaded below.
greedy_auc <- greedy_select_auc(d_filtered)
saveRDS(greedy_auc, here::here("data", "processed", "fits", "greedy_auc_results.rds"))
```


## Plot: person-average greedy selection

This plots the person-average selection path (Figure 10A): the solid line is the selected path, faint points are the non-selected candidates at each step, and the shaded band is the 95% bootstrap CI on the selected path. F is selected first and φ second, after which adding further parameters does not improve classification.

```{r fig.width=10, fig.height=6}
greedy_auc <- readRDS(here::here("data", "processed", "fits", "greedy_auc_results.rds"))

# Backward compatibility: rename alpha -> phi in cached results
if ("alpha" %in% names(greedy_auc$learner_fits)) setnames(greedy_auc$learner_fits, "alpha", "phi")
if ("param_added" %in% names(greedy_auc$learner_fits)) greedy_auc$learner_fits[param_added == "alpha", param_added := "phi"]
if ("param_added" %in% names(greedy_auc$summary))      greedy_auc$summary[param_added == "alpha", param_added := "phi"]
if ("free_params" %in% names(greedy_auc$summary)) {
  greedy_auc$summary[, free_params := lapply(free_params, function(fp) replace(fp, fp == "alpha", "phi"))]
}

summary_dt <- copy(greedy_auc$summary)
summary_dt[, param_label := param_display[param_added]]

selection_path <- summary_dt[selected == TRUE][order(step)]
selection_path[, param_label := param_display[param_added]]

seg_data <- summary_dt[selected == FALSE]
seg_data[, y_from := selection_path$auc_corrected[
  match(step - 1, selection_path$step)
]]

p_auc <- ggplot() +
  geom_segment(
    data = seg_data[!is.na(y_from)],
    aes(x = step - 1, xend = step, y = y_from, yend = auc_corrected,
        color = 'black'),
    linewidth = 0.4, linetype = "dashed", alpha = 0.7
  ) +
  geom_line(data = selection_path,
            aes(x = step, y = auc_corrected),
            linewidth = 1.2, color = "black") +
  geom_ribbon(data = selection_path,
              aes(x = step, ymin = auc_lo, ymax = auc_hi),
              alpha = 0.15, fill = "grey40") +
  geom_point(data = summary_dt[selected == FALSE],
             aes(x = step, y = auc_corrected, color = param_label),
             size = 3, shape = 16, alpha = 0.6) +
  geom_point(data = selection_path,
             aes(x = step, y = auc_corrected, color = param_label),
             size = 5, shape = 17) +
  geom_text(
    data = summary_dt[selected == FALSE & auc_corrected < 0.6],
    aes(x = step + 0.15, y = 0.62,
        label = paste0(param_label, " (",
                       scales::percent(auc_corrected, accuracy = 1), ")"),
        color = 'black'),
    size = 4, hjust = 0, show.legend = FALSE
  ) +
  scale_color_manual(name = "Parameter", values = param_cols) +
  scale_x_continuous(breaks = 1:5) +
  scale_y_continuous(limits = c(0.6, 1),
                     breaks = seq(0.6, 1, by = 0.05),
                     oob = scales::squish) +
  labs(x = "Step (parameter added)",
       y = "Classification AUC (bootstrap-corrected)") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none", panel.grid.minor = element_blank())

p_auc
```


## Single-session AUC from person-average selection path

Compute single-session AUCs for all candidate models fitted during the
person-average greedy selection (i.e., same fits, different evaluation criterion).

```{r}
summary_single <- map_dfr(
  split(greedy_auc$summary, seq_len(nrow(greedy_auc$summary))),
  function(s) {
    fp      <- s$free_params[[1]]
    step_i  <- s$step[[1]]
    param_i <- s$param_added[[1]]

    fits_i    <- greedy_auc$learner_fits[step == step_i & param_added == param_i]
    fits_i    <- merge(fits_i, clinical_df, by = "user_id")
    pred_cols <- unname(PARAM_COL[fp])
    pred_cols <- pred_cols[pred_cols %in% names(fits_i)]

    auc_res <- bootstrap_auc_corrected(
      fits_i$outcome, as.data.frame(fits_i[, ..pred_cols]), n_boot = 500
    )

    data.table(
      step = step_i, param_added = param_i, free_params = list(fp),
      auc_apparent = auc_res$auc_apparent, auc_corrected = auc_res$auc_corrected,
      auc_lo = auc_res$auc_lo, auc_hi = auc_res$auc_hi,
      selected = s$selected[[1]]
    )
  }
)

summary_single[, param_label := param_display[param_added]]
selection_single <- summary_single[selected == TRUE][order(step)]

seg_single <- summary_single[selected == FALSE]
seg_single[, y_from := selection_single$auc_corrected[
  match(step - 1, selection_single$step)
]]
```


## Greedy selection: single-session criterion (steps 3–5)

The selection path diverges from step 3 onward. Since steps 1–2 selected the
same parameters (F, then phi) under both criteria, we reuse those fits and
rerun only steps 3–5 with single-session AUC as the selection criterion.

```{r}
# ── Greedy selection from a given step (reusing earlier fits) ──────────────────
greedy_select_auc_from_step <- function(d_filtered,
                                        already_selected = c("F", "phi"),
                                        remaining_params = c("tau", "s", "ter"),
                                        start_step       = 3,
                                        n_boot           = 500,
                                        full_iter        = 100) {

  clinical_df <- unique(d_filtered[, .(user_id, clinical_status)])[
    , outcome := as.integer(clinical_status == "MCI")
  ]
  d_sessions <- split(d_filtered, by = "session_id")

  fit_sessions <- function(free_params) {
    message("  Fitting ", length(d_sessions), " sessions with free: {",
            paste(free_params, collapse = ", "), "} ...")
    results <- future_map(d_sessions, function(d_i) {
      fit_i <- fit_one_learner(d_i, free_params = free_params,
                               full_iter = full_iter)[.N]
      if (is.null(fit_i) || nrow(fit_i) == 0) return(data.table())
      fit_i[, `:=`(user_id         = d_i$user_id[1],
                   clinical_status = d_i$clinical_status[1],
                   session_id      = d_i$session_id[1])]
      fit_i
    }, .options = furrr_options(seed = TRUE), .progress = interactive())
    results <- results[!vapply(results, function(x) is.null(x) || nrow(x) == 0,
                               logical(1))]
    rbindlist(results, fill = TRUE)
  }

  selected  <- already_selected
  remaining <- remaining_params
  log_rows  <- list()
  all_fits  <- list()

  for (step in start_step:(start_step + length(remaining_params) - 1)) {
    message("\n=== Step ", step, ": testing ", length(remaining), " candidate(s) ===")
    step_rows <- list()

    for (candidate in remaining) {
      trial_params <- c(selected, candidate)
      message("  Candidate: {", paste(trial_params, collapse = ", "), "}")

      fits_dt <- fit_sessions(trial_params)
      if (nrow(fits_dt) == 0) next

      fits_with_outcome <- merge(fits_dt, clinical_df, by = "user_id")
      pred_cols <- unname(PARAM_COL[trial_params])
      pred_cols <- pred_cols[pred_cols %in% names(fits_with_outcome)]

      auc_res <- bootstrap_auc_corrected(
        fits_with_outcome$outcome,
        as.data.frame(fits_with_outcome[, ..pred_cols]),
        n_boot = n_boot
      )

      message(sprintf("    AUC corrected = %.3f [%.3f-%.3f]",
                      auc_res$auc_corrected, auc_res$auc_lo, auc_res$auc_hi))

      step_rows[[candidate]] <- data.table(
        step = step, param_added = candidate, free_params = list(trial_params),
        auc_apparent = auc_res$auc_apparent, auc_corrected = auc_res$auc_corrected,
        auc_lo = auc_res$auc_lo, auc_hi = auc_res$auc_hi, selected = FALSE
      )

      key      <- paste(step, candidate, sep = "_")
      fit_copy <- copy(fits_dt)
      fit_copy[, `:=`(step = step, param_added = candidate,
                      free_params_str = paste(trial_params, collapse = ","))]
      all_fits[[key]] <- fit_copy
    }

    step_dt  <- rbindlist(step_rows)
    best_idx <- which.max(step_dt$auc_corrected)
    best_par <- step_dt$param_added[best_idx]
    step_dt[best_idx, selected := TRUE]

    message(sprintf("  >>> Selected: %s  (AUC = %.3f)",
                    best_par, step_dt$auc_corrected[best_idx]))

    log_rows[[step - start_step + 1]] <- step_dt
    selected  <- c(selected, best_par)
    remaining <- setdiff(remaining, best_par)
  }

  list(summary = rbindlist(log_rows),
       learner_fits = rbindlist(all_fits, fill = TRUE))
}
```


```{r, eval=FALSE}
new_steps <- greedy_select_auc_from_step(
  d_filtered,
  already_selected = c("F", "phi"),
  remaining_params = c("tau", "s", "ter"),
  start_step       = 3,
  n_boot           = 500
)

# Recompute single-session AUC for reused steps 1–2
reused_summary <- map_dfr(
  split(greedy_auc$summary[step %in% 1:2],
        seq_len(nrow(greedy_auc$summary[step %in% 1:2]))),
  function(s) {
    fp      <- s$free_params[[1]]
    step_i  <- s$step[[1]]
    param_i <- s$param_added[[1]]

    fits_i    <- greedy_auc$learner_fits[step == step_i & param_added == param_i]
    fits_i    <- merge(fits_i, clinical_df, by = "user_id")
    pred_cols <- unname(PARAM_COL[fp])
    pred_cols <- pred_cols[pred_cols %in% names(fits_i)]

    auc_res <- bootstrap_auc_corrected(
      fits_i$outcome, as.data.frame(fits_i[, ..pred_cols]), n_boot = 500
    )

    data.table(
      step = step_i, param_added = param_i, free_params = list(fp),
      auc_apparent = auc_res$auc_apparent, auc_corrected = auc_res$auc_corrected,
      auc_lo = auc_res$auc_lo, auc_hi = auc_res$auc_hi,
      selected = s$selected[[1]]
    )
  }
)

greedy_auc_single <- list(
  summary      = rbind(reused_summary, new_steps$summary),
  learner_fits = rbind(greedy_auc$learner_fits[step %in% 1:2],
                       new_steps$learner_fits)
)

# Expensive; cached and reloaded in the next chunk.
saveRDS(greedy_auc_single, here::here("data", "processed", "fits", "greedy_auc_results_single.rds"))
```


## Plot: combined figure (person-average + single-session)

```{r fig.width=12, fig.height=6}
greedy_auc_single <- readRDS(here::here("data", "processed", "fits", "greedy_auc_results_single.rds"))

# Backward compatibility: rename alpha -> phi in cached results
if ("alpha" %in% names(greedy_auc_single$learner_fits)) setnames(greedy_auc_single$learner_fits, "alpha", "phi")
if ("param_added" %in% names(greedy_auc_single$learner_fits)) greedy_auc_single$learner_fits[param_added == "alpha", param_added := "phi"]
if ("param_added" %in% names(greedy_auc_single$summary))      greedy_auc_single$summary[param_added == "alpha", param_added := "phi"]
if ("free_params" %in% names(greedy_auc_single$summary)) {
  greedy_auc_single$summary[, free_params := lapply(free_params, function(fp) replace(fp, fp == "alpha", "phi"))]
}

# ── Panel B: single-session selection path ────────────────────────────────────
summary_single_b <- copy(greedy_auc_single$summary)
summary_single_b[, param_label := param_display[param_added]]

selection_single_b <- summary_single_b[selected == TRUE][order(step)]
selection_single_b[, param_label := param_display[param_added]]

seg_single_b <- summary_single_b[selected == FALSE]
seg_single_b[, y_from := selection_single_b$auc_corrected[
  match(step - 1, selection_single_b$step)
]]

p_auc_single <- ggplot() +
  geom_segment(
    data = seg_single_b[!is.na(y_from)],
    aes(x = step - 1, xend = step, y = y_from, yend = auc_corrected,
        color = param_label),
    linewidth = 0.4, linetype = "dashed", alpha = 0.7
  ) +
  geom_line(data = selection_single_b,
            aes(x = step, y = auc_corrected),
            linewidth = 1.2, color = "black") +
  geom_ribbon(data = selection_single_b,
              aes(x = step, ymin = auc_lo, ymax = auc_hi),
              alpha = 0.15, fill = "grey40") +
  geom_point(data = summary_single_b[selected == FALSE],
             aes(x = step, y = auc_corrected, color = param_label),
             size = 3, shape = 16, alpha = 0.6) +
  geom_point(data = selection_single_b,
             aes(x = step, y = auc_corrected, color = param_label),
             size = 5, shape = 17) +
  geom_text(
    data = summary_single_b[selected == FALSE & auc_corrected < 0.6] %>%
      arrange(auc_corrected) %>%
      mutate(y_pos = 0.62 + seq(0, by = 0.008, length.out = n())),
    aes(x = step + 0.15, y = y_pos,
        label = paste0(param_label, " (",
                       scales::percent(auc_corrected, accuracy = 1), ")"),
        color = 'black'),
    size = 4, hjust = 0, show.legend = FALSE
  ) +
  scale_color_manual(name = "Parameter", values = param_cols) +
  scale_x_continuous(breaks = 1:5) +
  scale_y_continuous(limits = c(0.6, 1),
                     breaks = seq(0.6, 1, by = 0.05),
                     oob = scales::squish) +
  labs(x = "Step (parameter added)",
       y = "Classification AUC (bootstrap-corrected)") +
  theme_bw(base_size = 12) +
  theme(legend.position = "right", panel.grid.minor = element_blank())

# change paramter order for the legend:
p_auc_single <- p_auc_single +
  scale_color_manual(name = "Parameter", values = param_cols,
                     breaks = names(param_cols))
# ── Combine panels ───────────────────────────────────────────────────────────
p_combined <- (p_auc + ggtitle("Participant average")) +
  (p_auc_single + ggtitle("Single session")) +
  plot_annotation(tag_levels = "A") & theme(plot.tag = element_text(size = 14, face = "bold"))

p_combined


# Paper Figure 10: greedy parameter selection by bootstrap-corrected AUC (A: participant average, B: single session)
ggsave(here::here("output", "greedy_selection_auc_combined.png"), p_combined,
       width = 11, height = 6, dpi = 300)
```


## Summary statistics for paper

Finally, we print the selection order and the AUCs of all candidates at each step, for both aggregation levels, as reported in Table 4 of the paper.

```{r}
cat("=== PANEL A: Person-average selection ===\n\n")
cat("Selection order:\n")
greedy_auc$summary[selected == TRUE][order(step),
  .(step, param_added,
    auc_corrected = round(auc_corrected, 3),
    auc_lo = round(auc_lo, 3),
    auc_hi = round(auc_hi, 3))] |> print()

cat("\nAll candidates at each step:\n")
greedy_auc$summary[order(step, -auc_corrected),
  .(step, param_added, selected,
    auc_corrected = round(auc_corrected, 3),
    auc_lo = round(auc_lo, 3),
    auc_hi = round(auc_hi, 3))] |> print()

cat("\n=== PANEL B: Single-session selection ===\n\n")
cat("Selection order:\n")
greedy_auc_single$summary[selected == TRUE][order(step),
  .(step, param_added,
    auc_corrected = round(auc_corrected, 3),
    auc_lo = round(auc_lo, 3),
    auc_hi = round(auc_hi, 3))] |> print()

cat("\nAll candidates at each step:\n")
greedy_auc_single$summary[order(step, -auc_corrected),
  .(step, param_added, selected,
    auc_corrected = round(auc_corrected, 3),
    auc_lo = round(auc_lo, 3),
    auc_hi = round(auc_hi, 3))] |> print()
```


# Session info

```{r}
sessionInfo()
```
